---
title: "Relationship Dissolution & Social Cognition Analysis"
author: "Millie Meng"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
params:
  use_simulated_data: true
---

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

required_packages <- c(
  "tidyverse", "janitor", "psych", "broom", "effectsize", "here", "knitr"
)
missing_packages <- required_packages[
  !vapply(required_packages, requireNamespace, logical(1), quietly = TRUE)
]
if (length(missing_packages) > 0) {
  stop("Install required packages first: ", paste(missing_packages, collapse = ", "))
}

library(tidyverse)
library(janitor)
library(psych)
library(broom)
library(effectsize)
library(here)
library(knitr)

dir.create(here("figures"), showWarnings = FALSE)
dir.create(here("results"), showWarnings = FALSE)

use_simulated_data <- isTRUE(params$use_simulated_data)
data_label <- if (use_simulated_data) "SIMULATED EXAMPLE" else "AUTHORIZED LOCAL DATA"
```

## Scope and reporting status

This notebook is a public analysis template. It contains no real participant-level data.

Confirmed study-level facts:

- survey **N = 119**;
- interviews **n = 19**;
- **KMO = .812**;
- **Cronbach's alpha = .891**; and
- ANOVA indicated a significant relationship between **attachment anxiety and rumination**.

Exact ANOVA statistics, effect sizes, regression estimates, factor loadings, and all other numerical results are **pending verification** and must not be inferred from the simulated demonstration.

## Data import or simulation

```{r data-import}
if (use_simulated_data) {
  set.seed(812)
  n <- 119

  # Demonstration structure only. Values are synthetic and are not calibrated to
  # reproduce the study's confirmed alpha, KMO, or inferential results.
  latent_attachment <- rnorm(n)
  demo_data <- tibble(
    participant_id = sprintf("SIM%03d", seq_len(n)),
    age = sample(18:55, n, replace = TRUE),
    attachment_group = sample(c("lower", "higher"), n, replace = TRUE),
    attachment_anxiety_1 = pmin(pmax(round(4 + latent_attachment + rnorm(n)), 1), 7),
    attachment_anxiety_2 = pmin(pmax(round(4 + latent_attachment + rnorm(n)), 1), 7),
    attachment_anxiety_3 = pmin(pmax(round(4 + latent_attachment + rnorm(n)), 1), 7),
    attachment_anxiety_4 = pmin(pmax(round(4 + latent_attachment + rnorm(n)), 1), 7),
    rumination_1 = sample(1:5, n, replace = TRUE),
    rumination_2 = sample(1:5, n, replace = TRUE),
    rumination_3 = sample(1:5, n, replace = TRUE),
    adjustment_score = round(rnorm(n, 50, 10), 1)
  )
} else {
  private_path <- here("data", "relationship_dissolution_data.csv")
  if (!file.exists(private_path)) {
    stop("Authorized local file not found. Keep private data in data/; it is git-ignored.")
  }
  demo_data <- readr::read_csv(private_path, show_col_types = FALSE)
}

raw_data <- demo_data |>
  clean_names()

cat("Analysis mode:", data_label, "\n")
cat("Rows loaded:", nrow(raw_data), "\n")
```

## Data validation and cleaning

```{r cleaning}
# Replace these names and allowed ranges with the final authorized codebook.
required_fields <- c(
  "participant_id", "attachment_group",
  paste0("attachment_anxiety_", 1:4),
  paste0("rumination_", 1:3), "adjustment_score"
)

missing_fields <- setdiff(required_fields, names(raw_data))
if (length(missing_fields) > 0) {
  stop("Missing required fields: ", paste(missing_fields, collapse = ", "))
}

clean_data <- raw_data |>
  distinct(participant_id, .keep_all = TRUE) |>
  mutate(
    attachment_group = factor(attachment_group),
    across(starts_with("attachment_anxiety_"), ~ na_if(.x, -99)),
    across(starts_with("rumination_"), ~ na_if(.x, -99))
  )

quality_summary <- tibble(
  metric = c("rows_after_deduplication", "duplicate_ids_removed", "total_missing_cells"),
  value = c(
    nrow(clean_data),
    nrow(raw_data) - nrow(clean_data),
    sum(is.na(clean_data))
  )
)
kable(quality_summary, caption = paste(data_label, "— quality checks"))
```

## Scale scoring

```{r scale-scoring}
# TODO before real-data use:
# 1. Confirm item membership.
# 2. Confirm reverse-keyed items and response ranges.
# 3. Define the minimum answered-item rule.
# Example reverse coding for a 1–7 item would be: reversed_item = 8 - original_item

scored_data <- clean_data |>
  mutate(
    attachment_anxiety = rowMeans(
      across(all_of(paste0("attachment_anxiety_", 1:4))), na.rm = TRUE
    ),
    rumination = rowMeans(
      across(all_of(paste0("rumination_", 1:3))), na.rm = TRUE
    )
  ) |>
  mutate(
    attachment_anxiety = if_else(is.nan(attachment_anxiety), NA_real_, attachment_anxiety),
    rumination = if_else(is.nan(rumination), NA_real_, rumination)
  )
```

## Reliability analysis

```{r reliability}
attachment_items <- scored_data |>
  select(all_of(paste0("attachment_anxiety_", 1:4)))

alpha_output <- psych::alpha(attachment_items, check.keys = FALSE)
alpha_table <- tibble(
  analysis = data_label,
  scale = "Attachment anxiety — example item set; verify",
  alpha = unname(alpha_output$total$raw_alpha),
  reporting_status = if (use_simulated_data) {
    "Simulated example; not a study result"
  } else {
    "Pending comparison with verified study output"
  }
)
kable(alpha_table, digits = 3)

# Confirmed study-level value (scale/item set must be checked in final records): alpha = .891.
```

## KMO and exploratory factor analysis

```{r kmo-efa}
psychometric_items <- scored_data |>
  select(starts_with("attachment_anxiety_"), starts_with("rumination_")) |>
  drop_na()

kmo_output <- psych::KMO(cor(psychometric_items))
kmo_table <- tibble(
  analysis = data_label,
  overall_msa = unname(kmo_output$MSA),
  reporting_status = if (use_simulated_data) {
    "Simulated example; not the confirmed study KMO"
  } else {
    "Pending comparison with verified study output"
  }
)
kable(kmo_table, digits = 3)

# Confirmed study-level KMO = .812.
# TODO: choose the number of factors using theory, scree/parallel analysis,
# and the final item set. The two-factor solution below is only a code example.
efa_example <- psych::fa(
  psychometric_items,
  nfactors = 2,
  rotate = "oblimin",
  fm = "minres"
)

efa_loadings <- as.data.frame(unclass(efa_example$loadings)) |>
  rownames_to_column("item")
head(efa_loadings) |> kable(digits = 3, caption = paste(data_label, "— example loadings"))
```

## ANOVA template

```{r anova}
# Placeholder operationalization: compares mean rumination across a categorical
# attachment group. Replace with the exact final model specification.
anova_model <- aov(rumination ~ attachment_group, data = scored_data)
anova_table <- broom::tidy(anova_model) |>
  mutate(
    analysis = data_label,
    reporting_status = if (use_simulated_data) {
      "Simulated example; do not report"
    } else {
      "Pending verification against final study output"
    }
  )
kable(anova_table, digits = 3)

eta_table <- effectsize::eta_squared(anova_model) |>
  as.data.frame() |>
  mutate(analysis = data_label)
kable(eta_table, digits = 3, caption = "Example effect-size workflow")

# Confirmed qualitative conclusion: the study ANOVA showed attachment anxiety
# and rumination were significantly related. Exact F, df, p, and effect size are
# not confirmed and therefore are not hard-coded here.
```

## Regression template

```{r regression}
# Example model only. Confirm the outcome, predictors, covariates, coding scheme,
# missing-data rule, and assumptions before running the final analysis.
regression_model <- lm(
  adjustment_score ~ attachment_anxiety + rumination + age,
  data = scored_data
)

regression_table <- broom::tidy(regression_model, conf.int = TRUE) |>
  mutate(
    analysis = data_label,
    reporting_status = if (use_simulated_data) {
      "Simulated example; do not report"
    } else {
      "Pending verification"
    }
  )
kable(regression_table, digits = 3)
```

## Visualization

```{r visualization, fig.width=7, fig.height=5}
relationship_plot <- scored_data |>
  ggplot(aes(x = attachment_anxiety, y = rumination, color = attachment_group)) +
  geom_point(alpha = 0.65) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    title = paste(data_label, "— attachment anxiety and rumination"),
    subtitle = "Demonstration graphic; not evidence of the study effect",
    x = "Attachment anxiety score (example scoring)",
    y = "Rumination score (example scoring)",
    color = "Attachment group"
  ) +
  theme_minimal(base_size = 12)

relationship_plot

ggsave(
  here("figures", "attachment-anxiety-rumination-example.png"),
  relationship_plot,
  width = 7,
  height = 5,
  dpi = 300
)
```

## Export results

```{r export}
readr::write_csv(alpha_table, here("results", "reliability-example.csv"))
readr::write_csv(kmo_table, here("results", "kmo-example.csv"))
readr::write_csv(anova_table, here("results", "anova-example.csv"))
readr::write_csv(regression_table, here("results", "regression-example.csv"))

writeLines(
  c(
    paste("Analysis mode:", data_label),
    "Public notebook outputs are simulated examples unless authorized real data were supplied.",
    "Confirmed study values: survey N=119; interviews n=19; KMO=.812; alpha=.891.",
    "Confirmed ANOVA conclusion: attachment anxiety and rumination were significantly related.",
    "All exact inferential statistics remain pending verification."
  ),
  here("results", "READ_ME_FIRST.txt")
)
```

## Session information

```{r session-info}
sessionInfo()
```
